---
title: "Fantasy Baseball Prep 2020"
author: "Seth Goldman"
date: "February 15, 2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(readr)
library(knitr)

all_rankings <- read_csv("./fantasy_2020/FantasyPros_Fantasy_Baseball_Rankings_ALL.csv")
```
```{r explore_all_rankings}
all_rankings %>% 
  head()

```

```{r explore_hitter_projs}
# Fantasy Pros Hitter Projections
proj_hitters <- read_csv("./fantasy_2020/FantasyPros_2020_Projections_H.csv")


# select scoring categories, and ABs
proj_hitters %>% 
  select(Player,Team,Positions,AB,R,HR,RBI,SB,OBP)


```

```{r explore_pitcher_projs}
# Fantasy Pros Pitcher Projections
proj_pitchers <- read_csv("./fantasy_2020/FantasyPros_2020_Projections_P.csv")

# select scoring categories (using IP as a proxy for QS)
proj_pitchers %>% 
  filter(Positions == "SP") %>% 
  select(Player, Team, Positions, K, ERA, WHIP, SV, IP)
  
# Create Correlelogram for SP scoring categories
sp_stat_correlations <- proj_pitchers %>%   
    filter(Positions == "SP") %>% 
  select(K, ERA, WHIP, IP) %>% 
  cor()

corrplot::corrplot(sp_stat_correlations,
                   type = "lower",
                   method = "number",
                   title = "SP Scoring Category Correlations - Projections")



# Create Correlelogram for RP scoring categories
rp_stat_correlations <- proj_pitchers %>%   
    filter(Positions == "RP") %>% 
  select(K, ERA, WHIP, IP, SV) %>% 
  cor()

corrplot::corrplot(rp_stat_correlations,
                   type = "lower",
                   method = "number",
                   title = "RP Scoring Category Correlations - Projections")


```

```{r explore 2019_season_results}

# Get season stats and calculate fantasy scoring based on (5x5 OBP / QS league)


batting_lahman <- read_csv("./fantasy_2020/Batting.csv",guess_max = 100000)

# Preview Hitter Scoring categories from 2018
batting_lahman %>% 
    filter(yearID == 2018,
           AB > 300) %>% 
  mutate("OBP" = round((H+BB+HBP)/(AB+BB+HBP+SF),3)) %>% 
  select(playerID, yearID, G, AB, R, HR, RBI,SB,OBP) %>% 
  View()


# Determine the TAM of each scoring category

batting_lahman %>% 
  filter(yearID == 2018,
           AB > 300) %>% 
  summarise("AB"=sum(AB),
            "H" = sum(H),
            "R" = sum(R),
            "HR" = sum(HR),
            "RBI" = sum(RBI),
            "SB" = sum(AB),
            "BB" = sum(BB),
            "HBP" = sum(HBP),
            "SF" = sum(SF)) %>% 
  mutate("PA" = AB+BB+HBP+SF,
         "RB" = H+BB+HBP) %>% 
  select(PA,R,HR,RBI,SB,RB)




batting_lahman %>% 
  filter(yearID > 2008,
           AB > 300) %>%
  group_by(yearID) %>% 
  summarise("AB"=sum(AB),
            "H" = sum(H),
            "R" = sum(R),
            "HR" = sum(HR),
            "RBI" = sum(RBI),
            "SB" = sum(AB),
            "BB" = sum(BB),
            "HBP" = sum(HBP),
            "SF" = sum(SF)) %>% 
  mutate("PA" = AB+BB+HBP+SF,
         "RB" = H+BB+HBP) %>% 
  select(yearID,PA,R,HR,RBI,SB,RB)

batting_lahman %>% 
  filter(yearID > 2012) %>% 
  group_by(yearID,playerID) %>% 
  summarise("AB"=sum(AB),
            "H" = sum(H),
            "R" = sum(R),
            "HR" = sum(HR),
            "RBI" = sum(RBI),
            "SB" = sum(SB),
            "BB" = sum(BB),
            "HBP" = sum(HBP),
            "SF" = sum(SF)) %>% 
  mutate("PA" = AB+BB+HBP+SF,
         "RB" = H+BB+HBP,
         "OBS" = round(RB/PA,3)) %>% 
  filter(PA > 400) %>%
  select(yearID,playerID, R, HR, RBI, SB, OBS) %>%  
  pivot_longer(cols = c("R","HR","RBI","SB","OBS"), 
    names_to = "metric", 
    values_to = "value") %>% ungroup() %>% 
  group_by(yearID, metric) %>% 
  top_n(189, value) %>%
  ungroup() %>% 
  group_by(yearID, metric) %>% 
  summarise("TAP" = sum(value)) %>% 
  ggplot(aes(x=yearID, y=TAP, color = metric))+
  geom_line()+
  facet_wrap(~metric)
  
# Need to not sum obp, and consider ballparking each cat based on 3 year trend



```

League Format:  
* Head-to-head matchups where you get a W for each of the 12 scoring category that you defeat your opponent, an L for each category you lose, and a Tie for the rest.  
* Six hitting categories: Runs, Runs Batted In, Stolen Bases, Home Runs, and On-base Percentage  
* Six pitching categories: Strikeouts, Quality Starts, Saves, ERA, and WHIP  

Rosters:  
* 26 players total  
* 21 starters (13 batters, 8 pitchers) 
* 5 bench positions  
* 2 slots for players on the Injured List, which do not count against the 26-player limit.  

Evaluating the Market:  
Since there are 12 teams in the league, and each team holds a roster of 26 players, there should be 312 rostered players at all times, plus a potential 24 players in the injury slots. For this reason, I only ought to be concerned with the 350(ish) most productive players.  

I also need to consider splitting batters and pitchers, since no one player (save for a couple exceptions), contribute to both scoring categories. From some quick research, I'm estimating that most 26-player rosters will hold between 14-15 batters and 11-12 pitchers. For the sake of estimation, I'll split the player pool 54/46 in favor of the batters. In other words, what I'll consider to be the top 350 players will be comprised of the top 189 batters, and the top 161 pitchers.


Total Production:  
To understand the relative value each player represents to my fantasy baseball team, I will try to estimate the total value generating by the top 350 players in a given year. which I'll call the Total Addressible Points (TAP). This will help identify the relative value of each rosterable player against the total market. 

Estimating the TAP:  
For each of the past few years, calculate each player's Fantasy Points, and calculate the sum of the top 350's scoring.


```{r fig.height=15}

fg_hitters <- read_csv("FanGraphs Leaderboard Hitters.csv")
fg_pitchers <- read_csv("FanGraphs Leaderboard Pitchers.csv")

fg_rankings <- bind_rows(fg_hitters,fg_pitchers) %>% 
  mutate(value = Dollars,
         "is_negative" = case_when(str_detect(string = value,
                                              pattern = fixed("($")) ~ 1,
                                   TRUE ~ 0),
         "remove_first" = str_remove(string = value,
                                     pattern = fixed("($")),
         "remove_second" = str_remove(string = remove_first,
                                      pattern =  fixed(")")),
         "remove_third" = as.numeric(str_remove(string = remove_second,
                                     pattern = fixed("$"))),
         "value" = ifelse(is_negative == 1, -1*remove_third, remove_third)) %>% 
  select(-remove_first,-remove_second,-remove_third)

rm(list=c("fg_hitters","fg_pitchers"))

# Select Top 350 most valuable players, plot distribution of value by position
fg_rankings %>% 
  filter(POS %in% c("SP","RP","OF","C","2B","SS","3B","1B")) %>% 
  top_n(value, n = 350) %>% 
  ggplot(aes(x=value))+
  geom_histogram()+
  facet_wrap(~POS)

# Summary of top 350 players by position with average value
fg_rankings %>% 
  filter(POS %in% c("SP","RP","OF","C","2B","SS","3B","1B")) %>% 
  top_n(value, n = 350) %>% 
  group_by(POS) %>% 
  summarize("Players" = n(),
            "Avg Value" = mean(value)) %>% 
  arrange(desc(Players)) %>% 
  kable()

fg_rankings %>% 
  filter(POS %in% c("SP","RP","OF","C","2B","SS","3B","1B")) %>% 
  group_by(POS) %>% 
  top_n(value,n = 20) %>%
  ungroup() %>% 
    ggplot(aes(x=reorder(PlayerName,value),y=value,color=POS))+
  geom_point()+
  coord_flip()+
  facet_wrap(~POS,scales = "free_y")+
  ggtitle("FanGraphs Dollar Value",
          subtitle = "Top 20 By Position")



```

